---
title: "Script_BEX_Paper"
author: "Mathis Nozais"
date: "04/11/2024"
output:
  html_document:
    code_folding: hide
    code_download: true
editor_options: 
  chunk_output_type: console
---

#################

Script for the mice scRNAseq for "BEX" paper.
Made for Docker SEURAT 440

#################

```{r}
library(Seurat)
library(biomaRt)
library(plotly)
library(tidyverse)
library(dplyr)
library(viridis)
library(patchwork) # to modify multi ggplot2 by Seurat
library(plotly)
library(ggrepel)
library(svglite)
library(paletteer)
library(dplyr)
library(reshape2)
```


Set your working directory to the folder clone from github
```{r}
# Set the variable to your working directory
WORKING_DIR <- "/yourcomputer/BEX/03_BulkRNA_analysis" # should point to the github folder that you clone
WORKING_DIR <- "/workspace/NASBioinfo/LALT/BIOINFO/BEX/GIT_TESTING/02_Seurat_analysis"  # to remove
#  Path to the folder that will contain output objects
OUTPUT_PATH <- (paste0(WORKING_DIR,"/03_Output/"))
```


# Pre-processing
From raw matrix (extracted from a bigger experiment with many other data)
```{r}
# loading the matrices
counts_matrix <- read.csv(paste0(WORKING_DIR,"/01_Data/RNA_matrix_sub.csv"), row.names = 1)
metadata <- read.csv(paste0(WORKING_DIR,"/01_Data/Metadata_matrix_sub.csv"), row.names = 1) 
ADT <- read.csv(paste0(WORKING_DIR,"/01_Data/ADT_matrix_sub.csv"), row.names = 1)
```

Creation of the Seurat object
```{r}
sub_Seurat <- CreateSeuratObject(counts = counts_matrix, assay = "RNA", meta.data = metadata)
```


```{r}
# Load in the UMI matrix
umi <- GetAssayData(object =sub_Seurat, slot = "counts")


#create an empty matrix containing NAs
Cell.list <- colnames(GetAssayData(object = sub_Seurat[["RNA"]], slot = "data" ) )
ADT.list <- c(unique(rownames(ADT)))
mat.adt <- matrix(nrow = length(ADT.list), ncol = length(Cell.list))
rownames(mat.adt) = ADT.list
colnames(mat.adt) = Cell.list

# Get cell barcodes detected by both RNA and ADT
joint_bcs <- intersect(colnames(umi),colnames(ADT))
adt <- as.matrix(ADT[,joint_bcs])

# Fill the empty matrix with values when existing
mat.adt[,joint_bcs]<-adt[,joint_bcs]

# Add ADT data as a new assay independent from RNA
sub_Seurat[["ADT"]] <- CreateAssayObject(counts = mat.adt[,colnames(sub_Seurat)])

# Normalize ADT data, here we use centered log-ratio (CLR) transformation
sub_Seurat <- NormalizeData(sub_Seurat, assay = "ADT", normalization.method = "CLR")

#Scale
sub_Seurat <- ScaleData(sub_Seurat, assay = "ADT")

```

```{r}
sub_Seurat <- NormalizeData(sub_Seurat,display.progress = FALSE) #testing that


sub_Seurat <- FindVariableFeatures(object = sub_Seurat,assay = "RNA", selection.method = "vst", nfeatures = 2000)

sub_Seurat  <- ScaleData(sub_Seurat,assay="RNA",verbose = FALSE, do.center = TRUE)
  
sub_Seurat <- RunPCA(object = sub_Seurat,
                    assay = "RNA",
                    verbose = FALSE, #if TRUE print the top genes for each PC
                    seed.use = 1234,
                    npcs = 50) # sur les 50 premieres composantes
  
ElbowPlot(sub_Seurat, ndims = 30, reduction = "pca")

DimPlot(sub_Seurat,reduction = "pca",dims = 9:10,group.by = "MULTI_ID") 
DimPlot(sub_Seurat,reduction = "pca",dims = 10:11,group.by = "MULTI_ID") 
DimPlot(sub_Seurat,reduction = "pca",dims = 11:12,group.by = "MULTI_ID") #nope
DimPlot(sub_Seurat,reduction = "pca",dims = 13:14,group.by = "MULTI_ID") # nope

sub_Seurat <- FindNeighbors(object = sub_Seurat, 
                           dims = 1:10, 
                           verbose = FALSE, 
                           force.recalc = TRUE, 
                           reduction = "pca")

DefaultAssay(sub_Seurat) <- "RNA"

sub_Seurat <- FindClusters(object = sub_Seurat, 
                          resolution =1 ,
                          verbose = FALSE,
                          random.seed = 1234)
sub_Seurat <- RunUMAP(object = sub_Seurat, reduction = "pca", seed.use = 1234, dims = 1:10)

DimPlot(sub_Seurat, group.by = "MULTI_ID")
```

## Cell cycle regression
```{r}
## Cell cycle
cc.genes.updated.2019 #for Human need to use bIOmart if using mice

human = useEnsembl(biomart = "ensembl", dataset = "hsapiens_gene_ensembl",version = 105) #Using a fixed version of the db as of DEc 21 allow to avoid getLDS error
mouse = useEnsembl(biomart = "ensembl", dataset = "mmusculus_gene_ensembl", version = 105)
genesV2 = getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values = cc.genes$s.genes , mart = human, attributesL = c("mgi_symbol"), martL = mouse, uniqueRows=T)
S.Genes_mouse <- unique(genesV2[, 2])

genesV2 = getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values = cc.genes$g2m.genes , mart = human, attributesL = c("mgi_symbol"), martL = mouse, uniqueRows=T)
G2M.Genes_mouse <- unique(genesV2[, 2])

#Assign Cell-Cycle Scores
sub_Seurat_reg <- CellCycleScoring(object = sub_Seurat, s.features = S.Genes_mouse, g2m.features = G2M.Genes_mouse, set.ident = TRUE)

 Tsne<-data.frame(
    tSNE_1 = sub_Seurat_reg@reductions$umap@cell.embeddings[,1],
    tSNE_2= sub_Seurat_reg@reductions$umap@cell.embeddings[,2],
    gene= sub_Seurat_reg@meta.data$S.Score
  )
  HTO= sub_Seurat_reg@meta.data$MULTI_ID
  Max=max(sub_Seurat_reg@meta.data$S.Score)
  Min=min(sub_Seurat_reg@meta.data$S.Score)
  ggplotly(ggplot(Tsne,aes(x=tSNE_1,y=tSNE_2))+geom_point(aes(color=gene,shape=HTO))+
    scale_colour_gradient2(low = "blue",mid="orange",high="red",name="S score",midpoint=(Max+Min)/2))
  
  Tsne<-data.frame(
    tSNE_1 = sub_Seurat_reg@reductions$umap@cell.embeddings[,1],
    tSNE_2= sub_Seurat_reg@reductions$umap@cell.embeddings[,2],
    gene= sub_Seurat_reg@meta.data$G2M.Score
  )
  HTO= sub_Seurat_reg@meta.data$MULTI_ID
  Max=max(sub_Seurat_reg@meta.data$G2M.Score)
  Min=min(sub_Seurat_reg@meta.data$G2M.Score)
  
  ggplotly(ggplot(Tsne,aes(x=tSNE_1,y=tSNE_2))+geom_point(aes(color=gene,shape=HTO))+
    scale_colour_gradient2(low = "blue",mid="orange",high="red",name="G2M score",midpoint=(Max+Min)/2))

  DimPlot(sub_Seurat_reg, reduction = "umap", group.by = "Phase")
```

Regression on the Seurat object
```{r}
#cell cycle regression
sub_Seurat_reg <- ScaleData(sub_Seurat_reg, vars.to.regress = c("S.Score", "G2M.Score"), features = sub_Seurat_reg@assays$RNA@var.features)
  
sub_Seurat_reg <- RunPCA(object = sub_Seurat_reg,
                    assay = "RNA",
                    verbose = FALSE, #if TRUE print the top genes for each PC
                    seed.use = 1234,
                    npcs = 50) #sur les 50 premieres composantes
```

### Re clustering without cycle
```{r}
ElbowPlot(sub_Seurat_reg, ndims = 30, reduction = "pca")

DimPlot(sub_Seurat_reg,reduction = "pca",dims = 7:8,group.by = "MULTI_ID") 
DimPlot(sub_Seurat_reg,reduction = "pca",dims = 8:9,group.by = "MULTI_ID") 
DimPlot(sub_Seurat_reg,reduction = "pca",dims = 9:10,group.by = "MULTI_ID") # 
DimPlot(sub_Seurat_reg,reduction = "pca",dims = 10:11,group.by = "MULTI_ID") #
DimPlot(sub_Seurat_reg,reduction = "pca",dims = 11:12,group.by = "MULTI_ID") # nope
DimPlot(sub_Seurat_reg,reduction = "pca",dims = 12:13,group.by = "MULTI_ID") # nope
DimPlot(sub_Seurat_reg,reduction = "pca",dims = 13:14,group.by = "MULTI_ID") # nope


sub_Seurat_reg <- FindNeighbors(object = sub_Seurat_reg, 
                           dims = 1:12, 
                           verbose = FALSE, 
                           force.recalc = TRUE, 
                           reduction = "pca")
DefaultAssay(sub_Seurat_reg) <- "RNA"
sub_Seurat_reg <- FindClusters(object = sub_Seurat_reg, 
                          resolution =1 ,
                          verbose = FALSE,
                          random.seed = 1234)
sub_Seurat_reg <- RunUMAP(object = sub_Seurat_reg, reduction = "pca", seed.use = 1234, dims = 1:12,min.dist = 0.5)

DimPlot(sub_Seurat_reg, group.by = "MULTI_ID")

save(sub_Seurat_reg, file=paste0(WORKING_DIR,"/03_Output/BC_BEX_mice.Robj"))
```

# Loading object
```{r}
load(file=paste0(WORKING_DIR,"/03_Output/BC_BEX_mice.Robj"))
```


```{r}
DimPlot(sub_Seurat_reg, group.by = "MULTI_ID", pt.size = 2)

## Fine tune the clustering
sub_Seurat_reg <- FindClusters(object = sub_Seurat_reg, 
                          resolution =0.6,
                          verbose = FALSE,
                          random.seed = 1234)

DimPlot(sub_Seurat_reg, group.by = "RNA_snn_res.0.6", pt.size = 2)
#Choose the clustering that separate main T cell population. Some clusters will be grouped together at the end.
```

## Rename mice
```{r}
Idents(sub_Seurat_reg) <- "MULTI_ID"
#Add tum_stage slot 
sub_Seurat_reg@meta.data$tum_stage <-  ""

# I'll call pre tum also physio to simplify
sub_Seurat_reg@meta.data[c(WhichCells(sub_Seurat_reg,idents = "Thymus-790")),]$tum_stage = "Tumoral"  
sub_Seurat_reg@meta.data[c(WhichCells(sub_Seurat_reg,idents = "Thymus-814")),]$tum_stage = "Physiological"  
sub_Seurat_reg@meta.data[c(WhichCells(sub_Seurat_reg,idents = "Thymus-813")),]$tum_stage = "Physiological"  

DimPlot(sub_Seurat_reg, group.by = "tum_stage", cols = c("#99C3FB","#ED7D31"),pt.size = 1.5)+ theme_void()
```

```{r}
#Plot for paper
image=DimPlot(sub_Seurat_reg, group.by = "tum_stage", cols = c("#99C3FB","#ED7D31"),pt.size = 1)+ theme_void()+NoLegend()

#This actually save the plot in a image
ggsave(file=paste0(WORKING_DIR,"/03_Output/Fig1_A.jpeg"), plot=image, width=2400, height=1600,units = "px",dpi = 320,device = "jpeg") # to test the ratio of image
ggsave(file=paste0(WORKING_DIR,"/03_Output/Fig1_A.svg"), plot=image, width=2400, height=1600,units = "px",dpi = 320, device = "svg")
```

## Population

### asign

```{r}
# Kept the smallest clustering to separate Dpblast and DN
DimPlot(sub_Seurat_reg, group.by = "RNA_snn_res.0.6", label = T,pt.size = 2, label.size = 20)

#fig paper
image <- DimPlot(sub_Seurat_reg, group.by = "RNA_snn_res.0.6",pt.size = 1,cols = paletteer_d("ggthemes::Tableau_10"))+ theme_void()+NoLegend()
ggsave(file=paste0(WORKING_DIR,"/03_Output/Fig1_C.svg"), plot=image, width=2400, height=1600,units = "px",dpi = 320, device = "svg")
```

```{r}
Idents(sub_Seurat_reg) <- "RNA_snn_res.0.6"  
sub_Seurat_reg@active.ident <- factor(sub_Seurat_reg@active.ident,levels=c("6","5","4","0","1","8","9","3","2","7"))

# detailed wo cluster merged
DotPlot(sub_Seurat_reg, features = c("adt_CD25","Ptcra","Sox13","Ikzf2","Trdc","Trac","Trbc2","Cd3d","Cd4","Cd8b1","Ccr9","Trp53inp1","Rag1","Sox4","Cd5","Cd27","Gzma","Hes1","Notch1","Top2a","Mki67","Cdk1","Cdk6","Myc"), dot.scale =5)+coord_flip()+
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_colour_gradient2(low = "blue3", mid = "white", high = "red") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))

## Fig paper
image <- DotPlot(sub_Seurat_reg, features = c("adt_CD25","Ptcra","Sox13","Ikzf2","Trdc","Trac","Trbc2","Cd3d","Cd4","Cd8b1","Ccr9","Trp53inp1","Rag1","Sox4","Cd5","Cd27","Gzma","Hes1","Notch1","Top2a","Mki67","Cdk1","Cdk6","Myc"), dot.scale =5)+coord_flip()+
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.7) +
  scale_colour_gradient2(low = "blue3", mid = "white", high = "red") +theme(axis.text=element_text(size=15),legend.text=element_text(size=15),axis.title=element_text(size=15))+
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))

#This actually save the plot in a image
ggsave(file=paste0(WORKING_DIR,"/03_Output/Fig1_B.jpeg"), plot=image, width=3200, height=2400,units = "px",dpi = 320,device = "jpeg")
ggsave(file=paste0(WORKING_DIR,"/03_Output/Fig1_B.svg"), plot=image, width=3200, height=2400,units = "px",dpi = 320,device = "svg")

## Asignn main cell population to group of cluster

sub_Seurat_reg$annotation <- ""

# extract cell bc per cluster
clus0 <- row.names(subset(sub_Seurat_reg@meta.data, RNA_snn_res.0.6 == "0"))
clus1 <- row.names(subset(sub_Seurat_reg@meta.data, RNA_snn_res.0.6 == "1"))
clus2 <- row.names(subset(sub_Seurat_reg@meta.data, RNA_snn_res.0.6 == "2"))
clus3 <- row.names(subset(sub_Seurat_reg@meta.data, RNA_snn_res.0.6 == "3"))
clus4 <- row.names(subset(sub_Seurat_reg@meta.data, RNA_snn_res.0.6 == "4"))
clus5 <- row.names(subset(sub_Seurat_reg@meta.data, RNA_snn_res.0.6 == "5"))
clus6 <- row.names(subset(sub_Seurat_reg@meta.data, RNA_snn_res.0.6 == "6"))
clus7 <- row.names(subset(sub_Seurat_reg@meta.data, RNA_snn_res.0.6 == "7"))
clus8 <- row.names(subset(sub_Seurat_reg@meta.data, RNA_snn_res.0.6 == "8"))
clus9 <- row.names(subset(sub_Seurat_reg@meta.data, RNA_snn_res.0.6 == "9"))


# Assignment
sub_Seurat_reg@meta.data[c(clus3,clus2,clus9,clus7),]$annotation = "DP_tum"  
sub_Seurat_reg@meta.data[c(clus4,clus0,clus1,clus8),]$annotation = "DP_Q"  
sub_Seurat_reg@meta.data[c(clus5),]$annotation = "DP_P"
sub_Seurat_reg@meta.data[c(clus6),]$annotation = "DN"


DimPlot(sub_Seurat_reg, group.by = "annotation",pt.size = 1.5)
```


## DGE
```{r}
# After the definition of the population will compare only tumoral cells to DP(P) as tumoral cell are cycling
######## DP(P) vs tum ###########
Idents(sub_Seurat_reg) <- "annotation"
DGE_Prolif <- FindMarkers(sub_Seurat_reg,ident.1 = "DP_tum",ident.2 = "DP_P")

#export to check gene manually
DGE_Prolif <- DGE_Prolif[order(DGE_Prolif$avg_log2FC,decreasing = TRUE),] # sort by log2fc
```

## Volcano
```{r}
##### Volcano plot #####
DGE_Prolif <- FindMarkers(sub_Seurat_reg,ident.1 = "DP_tum",ident.2 = "DP_P",min.pct = 0,logfc.threshold=0)

DGE_Prolif$abs_fold <- abs(DGE_Prolif$avg_log2FC)

DGE_Prolif$Significant <- ifelse(DGE_Prolif$p_val_adj < 0.05, ifelse(DGE_Prolif$abs_fold > 0.25 ,ifelse(DGE_Prolif$avg_log2FC >0, "Overexpressed in tumoral","Overexpressed in physio"),"Not Significative"), "Not Significative")



DGE_Prolif$genename <- rownames(DGE_Prolif) 
ggplot(DGE_Prolif, aes(x = avg_log2FC, y = -log10(p_val_adj))) +
  geom_point(aes(color = Significant))+
  guides(color = guide_legend(override.aes = list(size = 5)))+ #change point legend
   scale_color_manual(values = c("grey" ,"#4678BB","#9A4177")) +
   theme_bw(base_size = 12) + theme(legend.position = "bottom",text = element_text(size = 20 )) + #chnge all text size
  scale_x_continuous( limits=c(-4, 4))+
 geom_text_repel(
    data = subset(DGE_Prolif,abs_fold >1.3 & p_val_adj < 10^-40),
    aes(label = genename),
    size = 4,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines"),min.segment.length = 0)

```

## Pannel 1-D
```{r}
#figure paper
image <- ggplot(DGE_Prolif, aes(x = avg_log2FC, y = -log10(p_val_adj))) +
  geom_point(aes(color = Significant))+
  guides(color = guide_legend(override.aes = list(size = 5)))+ #change point legend
   scale_color_manual(values = c("grey" ,"#4678BB","#9A4177")) +
   theme_bw(base_size = 12) + theme(legend.position = "bottom",text = element_text(size = 15 )) + #chnge all text size
  scale_x_continuous( limits=c(-4, 4))+
 geom_text_repel(
    data = subset(DGE_Prolif,abs_fold >1.3 & p_val_adj < 10^-40),
    aes(label = genename),
    size = 4,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines"),min.segment.length = 0)


ggsave(file=paste0(WORKING_DIR,"/03_Output/Fig1_D.jpeg"), plot=image, width=2400, height=2400,units = "px",dpi = 320,device = "jpeg") # to test the ratio of image
ggsave(file=paste0(WORKING_DIR,"/03_Output/Fig1_D.svg"), plot=image, width=2400, height=2400,units = "px",dpi = 320, device = "svg")

```


```{r}
# Fig sup DGE T-ALL genes
DGE_Prolif$genename <- rownames(DGE_Prolif) 
image <- ggplot(DGE_Prolif, aes(x = avg_log2FC, y = -log10(p_val_adj))) +
  geom_point(aes(color = Significant))+
  guides(color = guide_legend(override.aes = list(size = 5)))+ #change point legend
   scale_color_manual(values = c("grey" ,"#4678BB","#9A4177")) +
   theme_bw(base_size = 12) + theme(legend.position = "bottom",text = element_text(size = 20 )) + #chnge all text size
  scale_x_continuous( limits=c(-4, 4))+
 geom_text_repel(
    data = subset(DGE_Prolif,abs_fold >1.3 & p_val_adj < 10^-40),
    aes(label = genename),
    size = 4,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines"),min.segment.length = 0)+
 geom_text_repel(
    data = DGE_Prolif[c("Myc","Notch1","Cdk6","Gata3","Il7r","Trp53","Ly6c1","Eno1b","Eno1","Ly6a"),],
    aes(label = genename, fontface = "bold.italic"),
    size = 4,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines"),min.segment.length = 0)

ggsave(file=paste0(WORKING_DIR,".03_Output/Figsup1_A.svg"), plot=image, width=2400, height=2400,units = "px",dpi = 320, device = "svg")
```


## BEX expression plot
```{r}
FeaturePlot(sub_Seurat_reg, features = c("Bex1","Bex2","Bex3","Bex4","Bex6"),pt.size = 1.5,cols =c("#F1EBE8","#E7ADAE","#E77CAE","#DB54A8","#BE2EB9","#7618CA","#5014BA","#230A6A")) + guide_area() + plot_layout(ncol = 3,guides = "collect")& theme_void() & theme(plot.title = element_text(hjust = 0.5, size = 30),legend.key.height = unit(1.6,'cm'), legend.key.width = unit(1,'cm'),legend.text=element_text(size=18))
```

```{r}
#FIGURE PAPER
image <- FeaturePlot(sub_Seurat_reg, features = c("Bex1","Bex2","Bex3","Bex4","Bex6"),pt.size = 1,cols =c("#F1EBE8","#E7ADAE","#E77CAE","#DB54A8","#BE2EB9","#7618CA","#5014BA","#230A6A")) + guide_area() + plot_layout(ncol = 3,guides = "collect")& theme_void() & theme(plot.title = element_text(hjust = 0.5, size = 15),legend.key.height = unit(1.6,'cm'), legend.key.width = unit(1,'cm'),legend.text=element_text(size=15))

ggsave(file=paste0(WORKING_DIR,"/03_Output/Fig1_E.svg"), plot=image, width=3000, height=2400,units = "px",dpi = 320,device = "jpeg") # to test the ratio of image
ggsave(file=paste0(WORKING_DIR,"/03_Output/Fig1_E.svg"), plot=image, width=3400, height=2400,units = "px",dpi = 320, device = "svg")

```

# Mice WT thymus

## B6
BEX_WT mice EMBO https://www.embopress.org/doi/full/10.15252/embj.2021110023
```{r}
load(file = paste0(WORKING_DIR,"/01_Data/ThymusWT_clean_cc-regressed_noADT.Robj")) #thymusWTclean.ccreg
thymusWTclean.ccreg_update = UpdateSeuratObject(object = thymusWTclean.ccreg)

thymusWTclean.ccreg_update <- FindClusters(object = thymusWTclean.ccreg_update, 
                          resolution =0.42 ,
                          verbose = FALSE,
                          random.seed = 1234)

DimPlot(thymusWTclean.ccreg_update, label = T)

# rename cluster based on suplementary figures "EV2"
Idents(thymusWTclean.ccreg_update) <- "RNA_snn_res.0.42"
thymusWTclean.ccreg_update@meta.data$regroupclusters <- "NA"
thymusWTclean.ccreg_update@meta.data[thymusWTclean.ccreg_update@meta.data$RNA_snn_res.0.42 == "8",]$regroupclusters <- "DP_samll"
thymusWTclean.ccreg_update@meta.data[thymusWTclean.ccreg_update@meta.data$RNA_snn_res.0.42 == "0",]$regroupclusters <- "DP_samll"
thymusWTclean.ccreg_update@meta.data[thymusWTclean.ccreg_update@meta.data$RNA_snn_res.0.42 == "1",]$regroupclusters <- "DN3a"
thymusWTclean.ccreg_update@meta.data[thymusWTclean.ccreg_update@meta.data$RNA_snn_res.0.42 == "2",]$regroupclusters <- "DN3b_DN4"
thymusWTclean.ccreg_update@meta.data[thymusWTclean.ccreg_update@meta.data$RNA_snn_res.0.42 == "3",]$regroupclusters <- "DP_CD69+"
thymusWTclean.ccreg_update@meta.data[thymusWTclean.ccreg_update@meta.data$RNA_snn_res.0.42 == "4",]$regroupclusters <- "DN1_DN2"
thymusWTclean.ccreg_update@meta.data[thymusWTclean.ccreg_update@meta.data$RNA_snn_res.0.42 == "5",]$regroupclusters <- "ISP_DPblast"
thymusWTclean.ccreg_update@meta.data[thymusWTclean.ccreg_update@meta.data$RNA_snn_res.0.42 == "6",]$regroupclusters <- "SP"
thymusWTclean.ccreg_update@meta.data[thymusWTclean.ccreg_update@meta.data$RNA_snn_res.0.42 == "7",]$regroupclusters <- "Tgd"

FeaturePlot(thymusWTclean.ccreg_update, features =c("Bex1","Bex2","Bex3","Bex4","Bex6"),ncol = 5)

Idents(thymusWTclean.ccreg_update) <- "regroupclusters"
thymusWTclean.ccreg_update@active.ident <- factor(thymusWTclean.ccreg_update@active.ident,
levels=c("DN1_DN2","DN3a","DN3b_DN4","ISP_DPblast","DP_samll","DP_CD69+","SP","Tgd"))


DotPlot(thymusWTclean.ccreg_update, features =c("Cd34","Kit","Cpa3","Il2ra","Ptcra","Tfrc","Mki67","Lat","Lck","Rag1","Cd8b1","Cd8a","Cd4","Cd69","Cd5","Trdc","Bex1","Bex2","Bex3","Bex4","Bex6"),dot.scale = 7)+
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_colour_gradient2(low = "blue3", mid = "white", high = "red") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))
```

```{r}
#Save plot Sup1
image <- DotPlot(thymusWTclean.ccreg_update, features =c("Cd34","Kit","Cpa3","Il2ra","Ptcra","Tfrc","Mki67","Lat","Lck","Rag1","Cd8b1","Cd8a","Cd4","Cd69","Cd5","Trdc","Bex1","Bex2","Bex3","Bex4","Bex6"),dot.scale = 7)+
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_colour_gradient2(low = "blue3", mid = "white", high = "red") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))+ theme(text=element_text(size=10), axis.text = element_text(size = 10))

ggsave(file=paste0(WORKING_DIR,"/03_Output/FigSup1_B.pdf"), plot=image, width=3600, height=1000,units = "px",dpi = 320, device = "pdf")
```


# Human WT thymus
## Physio paper 1
```{r}
#nature comm 6 sep 24
# 130 000 cells
#https://www.nature.com/articles/s41467-024-51767-y#data-availability

human_thym <- readRDS(paste0(WORKING_DIR,"/01_Data/thymus.sc.RDS"))
FeaturePlot(human_thym,c("BEX1","BEX2","BEX3","BEX4","BEX5","TAL1","TAL2"),cols =c("#F1EBE8","#E7ADAE","#E77CAE","#DB54A8","#BE2EB9","#7618CA","#5014BA","#230A6A"))
#VlnPlot(human_thym,c("BEX1","BEX3","BEX2","BEX5"))


#subset pop of interest
unique(human_thym$sub.class)


## re order
Idents(human_thym) <- "sub.class"  
human_thym@active.ident <- factor(human_thym@active.ident,
levels=c("DN_early","DN_re","DN_blast","DP_re","DP_blast","abT(entry)","CD8aa","CD8T","CD4T","CD8T_mem","CD4T_mem","Treg","Treg.diff","T_agonist","T_apoptosis","T_proliferating",
 "NKT","ILC3","B_naive","B_memory","B_trans","Plasma",
 "cTEC","mTEC","Fb","Fb_cycling","VSMC","Endo","Lymph",
 "Mono","Ery",
 "pDC","DC","Mac"))

DotPlot(human_thym, features =c("BEX1","BEX2","BEX3","BEX4","BEX5","TAL1","TAL2"),dot.scale = 7)+
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_colour_gradient2(low = "blue3", mid = "white", high = "red") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))

# export
image <- DotPlot(human_thym, features =c("BEX1","BEX2","BEX3","BEX4","BEX5","TAL1","TAL2"),dot.scale = 8)+
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_colour_gradient2(low = "blue3", mid = "white", high = "red") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))+ theme(text=element_text(size=11), axis.text = element_text(size = 11))


#This actually save the plot in a image
ggsave(file=paste0(WORKING_DIR,"/03_Output/FigSup_Human1.pdf"), plot=image, width=2000, height=3600,units = "px",dpi = 320, device = "pdf")

```

## Physio paper 2
PARK dataset converted 
```{r}
load(file = paste0(WORKING_DIR,"/01_Data/Park_dataset.robj"))

DimPlot(PARK_obj, group.by = "Anno_level_5")


FeaturePlot(PARK_obj, features = c("BEX1", "BEX2", "NGFRAP1", "BEX4", "BEX5", "TAL1", "TAL2"),cols =c("#F1EBE8","#E7ADAE","#E77CAE","#DB54A8","#BE2EB9","#7618CA","#5014BA","#230A6A"))

Idents(PARK_obj) <- "Anno_level_5"  
genes_of_interest <- c("BEX1", "BEX2", "NGFRAP1", "BEX4", "BEX5", "TAL1", "TAL2")
expr_data <- FetchData(PARK_obj, vars = genes_of_interest)
expr_data$sub.class <- Idents(PARK_obj)

library(dplyr)
mean_expr_by_class <- expr_data %>%
  group_by(sub.class) %>%
  summarise(across(all_of(genes_of_interest), mean))


heatmap_matrix <- as.matrix(mean_expr_by_class[, -1])  # Enlever la colonne 'sub.class'
rownames(heatmap_matrix) <- mean_expr_by_class$sub.class

heatmap_long <- melt(heatmap_matrix)
colnames(heatmap_long) <- c("Gene", "SubClass", "Expression")

ggplot(heatmap_long, aes(x = SubClass, y = Gene, fill = Expression)) +
  geom_tile(color = "white", size = 0.1) +  # Bordures blanches pour les cases
  scale_fill_viridis(name = "Expression", option = "magma") +  # Palette viridis
  coord_equal() +  # Pour garder les cases carrées
  theme_minimal() +  # Style épuré
  ggtitle("Heatmap des moyennes d'expression par type cellulaire") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Rotation des labels
        axis.title.x = element_blank(),  # Pas de titre d'axe x
        axis.title.y = element_blank())  # Pas de titre d'axe y


# Set a specific order of annotation level
unique(PARK_obj$Anno_level_5)
unique(PARK_obj$Anno_level_4) # maybe easier to read
Idents(PARK_obj) <- "Anno_level_4"  




PARK_obj@active.ident <- factor(PARK_obj@active.ident,
levels=c("ETP","DN","DP","αβT(entry)","CD8αα(I)","CD8αα(II)",'γδT',"CD8+T","CD4+T","CD8+Tmem",'CD4+Tmem',"Th17","Treg","Treg(diff)","T(agonist)"
  ,"NKT","NK","ILC3","B_naive","B_memory","B_pro/pre","B_plasma",
  "mTEC","cTEC","TEC(neuro)","TEC(myo)","Epi_GCM2","Fb_1","Fb_2","Fb_cycling","VSMC","Endo","Lymph","NMP",
"Mono","Mgk",'Ery',"Mast",
"pDC","DC1","DC2","aDC","Mac"
)
)

"Lymph" # lylphatic endothelial cells
"NMP" # don't know where to put it

# T lymphoid, Thymic microevn, B Lympoid, other lymph, Myeloid, DC

# Dot plot fig sup
image <- DotPlot(PARK_obj, features =c("BEX1","BEX2","NGFRAP1","BEX4","BEX5","TAL1","TAL2"),dot.scale = 8)+
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_colour_gradient2(low = "blue3", mid = "white", high = "red") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))+  theme(text=element_text(size=11), axis.text = element_text(size = 11))


#This actually save the plot in a image
ggsave(file=paste0(WORKING_DIR,"/03_Output/FigSup_Human2.pdf"), plot=image, width=2000, height=3600,units = "px",dpi = 320, device = "pdf")


# Dot plot version fig principal, without too much detail
PARK_obj # subset park dataset to keep only subtype of interest for major fig
PARK_sub <- subset(x = PARK_obj, idents = c("ETP","DN","DP","αβT(entry)","CD8αα(I)","CD8αα(II)",'γδT',"CD8+T","CD4+T","CD8+Tmem",'CD4+Tmem',"Th17","Treg","Treg(diff)","T(agonist)","NKT","NK", "mTEC","cTEC"))

image <- DotPlot(PARK_sub, features =c("BEX1","BEX2","NGFRAP1","BEX4","BEX5","TAL1","TAL2"),dot.scale = 8)+
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.5) +
  scale_colour_gradient2(low = "blue3", mid = "white", high = "red") +
  guides(size=guide_legend(override.aes=list(shape=21, colour="black", fill="white")))+ coord_flip()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+  theme(text=element_text(size=11), axis.text = element_text(size = 11))

ggsave(file=paste0(WORKING_DIR,"/03_Output/Fig2_C.pdf"), plot=image, width=4000, height=1000,units = "px",dpi = 320, device = "pdf")
```





