---
output: html_document
author: "Delphine Potier"
output:
  html_document:
    code_folding: hide
    code_download: true
editor_options:
  chunk_output_type: console
---

#################

Script for Cut & Tag analysis of H3K27Ac mark in Jurkat WT and Jurkat CRISPR-edited clones (1D9,2G5 and 1B6).
Made with Docker rna431-2 (available on Zenodo)

#################

```{r}
library(DESeq2)
library(ggplot2)
library(Rsubread)
library(clusterProfiler)
library(dplyr)
library(ComplexHeatmap)
library(ggrepel)
library(RColorBrewer)
library(pathview)
library(DT)
```

H3K27Ac mark from Jurkat clone (BEX WT) will be compared to 3 others 1D9, 2G5 & 3B6 (that are BEX KO)

```{r}
# Set the variable to your working directory
WORKING_DIR <- "BEX/" # to be adapted
#  Path to the folder that will contain output objects
OUTPUT_PATH <- (paste0(WORKING_DIR,"03_Output/"))
```

#import feature count data
```{r}
count_mtx<-read.table(file = paste0(WORKING_DIR,"03_Output/H3K4me3_H3K27ac_narrow_nomodel/bwa/merged_library/macs3/narrow_peak/consensus/K27Ac/K27Ac.consensus_peaks.featureCounts.txt"),header = T)
head(count_mtx) # Check the names order to shorter names below

# Prep count matrix for DESeq2
rownames(count_mtx) <- paste(count_mtx$Geneid,count_mtx$Chr,count_mtx$Start,count_mtx$End,sep = "_")
count_mtx <- as.data.frame(count_mtx[,-c(1:6)]) #exclude non informative column
names(count_mtx) <- c("2G5_K27Ac","3B6_K27Ac","1D9_K27Ac","JKT_K27Ac") #shorter names

#preparation of sample info matrix
condition <- factor(c("BexNeg","BexNeg","BexNeg","BexPos"))
colData <- data.frame(row.names = colnames(count_mtx), condition)
```


# Cleaning data
```{r}
DESeq.ds <- DESeqDataSetFromMatrix(countData = count_mtx, colData = colData, design = ~ condition)
dds <- DESeq(DESeq.ds)
#saveRDS(dds, file = paste0(WORKING_DIR,"03_Output/dds_K27Ac_BexPos_vs_Bexneg.rds"))
```

# Loading Deseq obj
```{r}
dds <- readRDS(file = paste0(WORKING_DIR,"03_Output/dds_K27Ac_BexPos_vs_Bexneg.rds"))

# Extract count for GEO/SRA
raw_counts <- as.data.frame(counts(dds, normalized=FALSE))
raw_counts <- cbind(Region_ID = rownames(raw_counts), raw_counts)
write.table(raw_counts, file=paste0(WORKING_DIR,"03_Output/H3K27ac_raw_counts.txt"),sep = "\t", quote = FALSE, row.names = FALSE) #available in GEO
norm_count <- counts(dds, normalized=TRUE)
norm_count <- cbind(Region_ID = rownames(norm_count), norm_count)
write.table(norm_count, file=paste0(WORKING_DIR,"03_Output/H3K27ac_normalized_counts.txt"),sep = "\t", quote = FALSE, row.names = FALSE) #available in GEO
```


# DGE Bex- VS Bex+
```{r}
res <- results(dds, contrast = c("condition", "BexNeg", "BexPos"))
res_sorted_by_padj <- res[order(res$padj), ]
write.table(res_sorted_by_padj, file=paste0(WORKING_DIR,"03_Output/dds_H3K27Ac_BexPos_vs_Bexneg_results.txt"),sep = "\t")
```

```{bash}
# ### Bed files for i-cisTarget
# cat 03_Output/dds_H3K27Ac_BexPos_vs_Bexneg_results.txt | grep -v baseMean | awk '$3 < -1 && $7 < 0.05 {print $1}'| sed s/\"//g | awk -F'_' '{s=""; for(i=3;i<=NF-2;i++) s=s $i OFS; sub(OFS "$","",s); print s" "$(NF-1)" "$NF" "$1"_"$2}' OFS="_" >03_Output/H3K27Ac_down_Padj-0.05_Log2FC_1.bed
# cat 03_Output/dds_H3K27Ac_BexPos_vs_Bexneg_results.txt | grep -v baseMean | awk '$3 > 1 && $7 < 0.05 {print $1}'| sed s/\"//g | awk -F'_' '{s=""; for(i=3;i<=NF-2;i++) s=s $i OFS; sub(OFS "$","",s); print s" "$(NF-1)" "$NF" "$1"_"$2}' OFS="_" >03_Output/H3K27Ac_up_Padj-0.05_Log2FC_-1.bed
``

## Volcano
```{r}
df <- as.data.frame(res_sorted_by_padj)
df$absfoldchange <- abs(df$log2FoldChange)
df$regionname <- rownames(df)

df$Significant <- ifelse(df$padj < 0.05 & df$absfoldchange >1, "Significative","Not Significative")
dim(na.omit(df[df$Significant == "Significative",]))

size_axis <- max(abs(ceiling(min(na.omit(df$log2FoldChange)))), abs(ceiling(max(na.omit(df$log2FoldChange))))) #Get the extreme value of foldchange to set automaticaly limit of the ggplot and scale it around 0

#### Volcano plot #####
ggplot(df, aes(x = log2FoldChange, y = -log10(padj))) +
  geom_point(aes(color = Significant))+
   scale_color_manual(values = c("#3A3A3A","#C00000" ,"#3A3A3A")) +
   theme_bw(base_size = 12) + theme(legend.position = "bottom")  +geom_vline(xintercept=c(-1, 1), col="black")+
        geom_hline(yintercept=-log10(0.05), col="black")+
 geom_text_repel(
    data = subset(df, df$padj < 10^-4),
    aes(label = regionname),
    size = 4,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines"))+
  xlim(-(size_axis), size_axis)

plotMA(res, ylim = c(-4, 4))
```

```{r}
# Get number of differentially expressed genes
dim(na.omit(df[df$Significant == "Significative",]))
dim(na.omit(df[df$Significant == 'Significative' & df$log2FoldChange > 1, ]))
dim(na.omit(df[df$Significant == 'Significative' & df$log2FoldChange < 1, ]))
```


## Heatmap
```{r}
vsd <- vst(dds,blind = FALSE)

mat_vst <- assay(vsd)
vst_df <- as.data.frame(mat_vst)
vst_df <- cbind(Region_ID = rownames(vst_df), vst_df)
write.table(vst_df, file=paste0(WORKING_DIR,"03_Output/H3K27ac_vst_counts.txt"), sep = "\t", quote = FALSE, row.names = FALSE) #available in GEO

region_vec <- na.omit(df[df$Significant == "Significative",])$regionname
mat_vst_sub <- mat_vst[region_vec,]
scaled_mat = t(scale(t(mat_vst_sub))) # row centered matrix Z score
hm_colors = c("dodgerblue1","white","red2")

#to automate scale
hm_limit =c(min(scaled_mat),0, max(scaled_mat))

column_order <- c("JKT_K27Ac","1D9_K27Ac","2G5_K27Ac","3B6_K27Ac")
scaled_mat <- scaled_mat[, column_order]
annotation_df <- colData
annotation_df <- annotation_df[column_order, , drop = FALSE]
column_annotation <- HeatmapAnnotation(
  df = annotation_df,
  col = list(condition = c("BexNeg" = "blue","BexPos"="black"))
)

set.seed(123) # for kmean reproducibility
ComplexHeatmap::Heatmap(scaled_mat,show_column_names = T ,cluster_columns = F,cluster_rows = T,column_names_side = "top",top_annotation = column_annotation, row_split = 2, column_split = annotation_df$condition)
```

# Sample distances
```{r distance_heatmap, fig.height = 4, fig.width = 10, results = 'asis'}
############# Heatmap hclsut(dist) ############
print("calculated on all regions")

sampleDists <- dist(t(assay(vsd)))
sampleDistMatrix <- as.matrix(sampleDists)
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)

my_sample_col <- colnames(assay(vsd))
rownames(sampleDistMatrix) <- my_sample_col
colnames(sampleDistMatrix) <- my_sample_col
pheatmap(sampleDistMatrix, clustering_distance_rows=sampleDists, clustering_distance_cols=sampleDists,col=colors)


var_regions <- apply(assay(vsd), 1, var)
var_regions <- sort(var_regions,decreasing=TRUE)

print("calculated on top 500 ∆ regions")
top500_var_regions=names(sort(var_regions,decreasing=TRUE)[1:500])
highly_variable_regs <- vsd[top500_var_regions,]
sampleDists <- dist(t(assay(highly_variable_regs)))
sampleDistMatrix <- as.matrix(sampleDists)
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)

rownames(sampleDistMatrix) <- my_sample_col
colnames(sampleDistMatrix) <- my_sample_col
pheatmap(sampleDistMatrix, clustering_distance_rows=sampleDists, clustering_distance_cols=sampleDists,col=colors)
```
